Country-Level Impact of Education on Entrepreneurship

### 2014-12-12 PresentationandPaper/FromEducationToEntrepreneurship.Rmd
minor edit

---
title: "Country-Level Impact of Education on Entrepreneurship"
author: "Sam Dundon and Thierry Racaud"
date: "10 December 2014"
linestretch: 2 
output: pdf_document
bibliography: Main.bib
---

@@

Whilst it is tempting to look at education systems in a very pragmatic way with ratios of teachers per student, or number of students per class, we suggest to use another kind of indicator in order to trace the potential long-term investment made by governments. Indeed, we believe that education systems can favour entrepreneurship and, beyond, GDP growth at the country-level, depending on how supportive of students' self-confidence and how socially inclusive they are. We see this process as many-fold but will only focus here on one of the potential causal relations. We want to understand specifically to want extent the support of individual self-confidence within education systems can affect the likelihood to become an entrepreneur, and thus impact the rate of entrepreneurship at the country level.

@@

Grilo and Thurik [@Grilo2005b] who looked into what determines entrepreneurship engagement levels throughout the European Union and the US pointed especially at the level of financial support available for entrepreneurs in the country and also at administrative complexities.

Of course, what determines entrepreneurship at a country-level has been long searched already but although structural and environmental factors have been looked at extensively, less has been done regarding cultural or educational factors. Yet, “the relative stability of differences in entrepreneurship across countries suggests that factors other than economic ones are at play” [@freytag2010entrepreneurship]. Even if “in recent years research has increasingly devoted itself to the subject” [@OECD2009], a lot still needs to be done.

Over a hundred years have passed since Max Weber tried to explain how religious-ethical motivations affect entrepreneurship [@weberprotestant], and today religion may be less effective in justifying why some are more likely to start businesses than others. Still, that was a founding stone and it is now widely admitted that “cultural features influence attitudes towards start-ups and that these attitudes, in turn, have an impact on start-up activities” [@OECD2009].

More recently, Davidsson, along with Delmar at first (1992), then alone (1995), then with Wiklund [@davidsson1997values], at his turn, laid a stepping stone by investigating the relationship between structural factors, mainly economic, and cultural factors. As they compared two Swedish regions similar in terms of structure and observed different outcomes, they too concluded that there might be a small cultural effect in attitudes toward start-up activities. In 2010, Freytag and Thurik [@freytag2010entrepreneurship] have also studied determinants of entrepreneurship in a cross-country setting and pointed at cultural features. Their results showed that "country specific (cultural) variables seem to explain the preference for entrepreneurship" and self-employment. Yet, not so many empirical studies have been carried out on this topic and most of them focused on which attitudes were favourable to starting a business, while almost none explored what determined these dispositions.

Rather than trying to trace cultural features at large, as we mentioned earlier, we intend to focus specifically on one key cultural institution, that is the education system. Indeed, like many in the education literature tend to assert, we believe that school personnel have an affect on students' self-confidence [@Scott1996] which is key as "individuals who decide to start a new business exhibit more confidence than those who do not" [@hogarth2012entrepreneurial]. Now in entrepreneurship literature, "self-confidence was found to depend almost exclusively on the individual features of the respective person and his or her integration in social networks " [@OECD2009].

Still, if we look at the following data from the [Global Entrepreneurship Monitor (GEM)](http://www.gemconsortium.org) Adult Population Survey (APS) from 2012, there seem to be different amounts of self-confidence and entrepreneurial self-efficacy between countries. By a country's entrepreneurial self-efficacy we mean the belief that the average individual of that country holds about his or her abilities to succeed as an entrepreneur. Here, the variable "perceived capabilities" has to be understood in a narrow sense, it refers only to the perceived capabilities to start a business. Seemingly the variable "fear of failure" only refers to the fear of failure as an obstacle in starting a business, despite the perceived good opportunities.

@@

1 Beliefs  
2 Attitudes  
3 Intentions  
4 Behaviour

There could even be an incidence of the intention to perform a certain behaviour on the actual performance of that behaviour, which could lead to look into skills, abilities and willpower. It is almost like "individuals would be predisposed to entrepreneurial intentions based upon a combination of both personal and contextual factors" [@boyd1994influence] but Boyd and Vosikis building upon social psychology research introduce the concept of self-efficacy in Bird's model of entrepreneurial intentionality for "self-efficacy will moderate the relationship between the development of entrepreneurial intentions and the likelihood that these intentions will result in entrepreneurial actions or behaviour". Therefore, we will have to control for this entrepreneurial self-efficacy as we research potential effects of positive reinforcement.

Eventually, this goes without saying, we also looked into other studies connecting educational features and entrepreneurship. As a matter of fact, most of the research tends to focus specifically on entrepreneurial education, whatever its form, as it is expected partly to support the entrepreneurial self-efficacy [@boyd1994influence]. One slightly different and interesting perspective is that of King and Sobel [@Kerry2008], who have already tried to show how school choice, and thus a different kind of interactions with students, can increase the rate of youth entrepreneurship but their perspective was still not directly in line with our research design, they  asserted that a more business-like environment in K-12 education was the key, which was favoured by voucher programmes.

@@

At this stage, having explained how determinants of entrepreneurship have been looked into, having mentioned how education has been related to entrepreneurship so far, we need to assert here our very research question which is:

To what extent does the support of students' self-confidence within the education system impacts the rate of entrepreneurship across OECD countries?

@@

---> Education system reinforcing self-confidence

---> Higher degree of individual self-confidence

---> Higher probability an individual will become an entrepreneur

---> higher rate of entrepreneurship at a country-level.

@@

# Import csv of df with imputed figures (created in Index_sources/Amelia_source)
CI <- repmis::source_data("https://raw.githubusercontent.com/SamDund/Assign2Education/gh-pages/GEMData/Cimputed5.csv", header = TRUE, sep = ",", na.strings = NA)

>>>>

---
title: "Country-Level Impact of Education on Entrepreneurship"
author: "Sam Dundon and Thierry Racaud"
date: "12 December 2014"
linestretch: 2 
output: pdf_document
bibliography: Main.bib
---

@@

While it is tempting to look at education systems in a very pragmatic way with ratios of teachers per student, or number of students per class, we suggest to use another kind of indicator in order to trace the potential long-term investment made by governments. Indeed, we believe that education systems can favour entrepreneurship and, beyond, GDP growth at the country-level, depending on how supportive of students' self-confidence and how socially inclusive they are. We see this process as many-fold but will only focus here on one of the potential causal relations. We want to understand specifically to want extent the support of individual self-confidence within education systems can affect the likelihood to become an entrepreneur, and thus impact the rate of entrepreneurship at the country level.

@@

This framework and studies such as that by Grilo and Thurik [@Grilo2005b], who looked into what determines entrepreneurship engagement levels throughout the European Union and the U.S., provides a helpful starting point for our analysis.  However, rather than focusing on the structural and environmental factors driving the rate of entrepreneurship at a country-level -- which as been researched extensively already -- we have chosen to examine cultural and educational factors. We reached this decision following research indicating that “the relative stability of differences in entrepreneurship across countries suggests that factors other than economic ones are at play” [@freytag2010entrepreneurship]. Even if “in recent years research has increasingly devoted itself to the subject” [@OECD2009], a lot still needs to be done.

Research on these factors began as early as a hundred years ago, when Max Weber attempted to explain how religious-ethical motivations affect entrepreneurship [@weberprotestant].  Today, religion may be less effective in justifying why some are more likely to start businesses than others, but this effort served as a founding stone and it is now widely recognized that “cultural features influence attitudes towards start-ups and that these attitudes, in turn, have an impact on start-up activities” [@OECD2009].

More recently, Davidsson, along with Delmar at first (1992), then alone (1995), then with Wiklund [@davidsson1997values], at his turn, laid subsequent stepping stones by investigating the relationship between structural factors, mainly economic, and cultural factors. As they compared two Swedish regions similar in terms of structure and observed different outcomes, they too concluded that there might be a small cultural effect in attitudes toward start-up activities. In 2010, Freytag and Thurik [@freytag2010entrepreneurship] have also studied determinants of entrepreneurship in a cross-country setting and pointed at cultural features. Their results showed that "country specific (cultural) variables seem to explain the preference for entrepreneurship" and self-employment. Yet, not so many empirical studies have been carried out on this topic and most of them focused on which attitudes were favourable to starting a business, while almost none explored what determined these dispositions.

Rather than trying to trace cultural features at large, as we mentioned earlier, we intend to focus specifically on one key cultural institution, the education system. Indeed, like many in the education literature tend to assert, we believe that school personnel have an affect on students' self-confidence [@Scott1996], which is key as "individuals who decide to start a new business exhibit more confidence than those who do not" [@hogarth2012entrepreneurial]. Now in entrepreneurship literature, "self-confidence was found to depend almost exclusively on the individual features of the respective person and his or her integration in social networks " [@OECD2009].

Still, if we look at the following data from the [Global Entrepreneurship Monitor (GEM)](http://www.gemconsortium.org) Adult Population Survey (APS) from 2012, there seem to be different degrees of self-confidence and entrepreneurial self-efficacy between countries. By a country's entrepreneurial self-efficacy we mean the belief that the average individual of that country holds about his or her abilities to succeed as an entrepreneur. Here, the variable "Perceived Capabilities" should be understood in a narrow sense, it refers only to the perceived capabilities to start a business. Seemingly the variable "Fear of Failure" only refers to the "Fear of Failure" as an obstacle in starting a business, despite the "Perceived Opportunities."

@@

\begin{singlespace}

1. Beliefs

2. Attitudes

3. Intentions

4. Behaviour  
\end{singlespace}

There could even be an incidence of the intention to perform a certain behaviour on the actual performance of that behaviour, which could make research into skills, abilities and willpower an important step in understanding entrepreneurship rates. This sentiment was echoed by Boyd, who stated "individuals would be predisposed to entrepreneurial intentions based upon a combination of both personal and contextual factors" [@boyd1994influence]. This has been further stregthened in research by Boyd and Vosikis, building upon social psychology research, as they introduced the concept of self-efficacy in Bird's model of entrepreneurial intentionality, saying "self-efficacy will moderate the relationship between the development of entrepreneurial intentions and the likelihood that these intentions will result in entrepreneurial actions or behaviour." Therefore, we will need to control for this entrepreneurial self-efficacy as we research the potential effects of positive reinforcement.

We also looked into other studies connecting educational features and entrepreneurship. In fact, most of the research tends to focus specifically on entrepreneurial education, whatever its form, as it is expected partly to support the entrepreneurial self-efficacy [@boyd1994influence]. One slightly different and interesting perspective is that of King and Sobel [@Kerry2008], who have already tried to show how school choice, and thus a different kind of interactions with students, can increase the rate of youth entrepreneurship.  However, their perspective was still not directly in line with our research design, as they  asserted that a more business-like environment -- favoured by voucher programs -- in K-12 education was the key answer to their research question.

@@

At this stage, having explained how the determinants of entrepreneurship have been researched and having mentioned how education has been related to entrepreneurship so far, we need to assert here our research question:

*To what extent does the support of students' self-confidence within the education system impacts the rate of entrepreneurship across OECD countries?*

@@

1. An education system reinforcing self-confidence

2. Higher degree of individual self-confidence

3. Higher probability an individual will become an entrepreneur

4. Higher rate of entrepreneurship at a country-level

@@

# Import csv of df with imputed figures (created in Index_sources/Amelia_source)
CI <- repmis::source_data("https://raw.githubusercontent.com/SamDund/Assign2Education/gh-pages/MiscData/GEMData/Cimputed5.csv", header = TRUE, sep = ",", na.strings = NA)
